---
title: "yumCoffee"
output:
  html_document:
    fig_width: 8
    fig_height: 5
geometry: margin=0.75in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a project I decided to take on after I finished my first statistical computing class. I was on a coding high then, and I'm 99% confident that the high is still going.

In this project, I collected my own tasting profile for coffee and decided to plot my patterns. It's funny because as I did this plotting project and as I continue to collect more data, I can see some obvious patterns in my data that I actually didn't notice of myself.

```{r, include=FALSE, echo=FALSE}
# LIBRARIES
library(stringr)
library(ggplot2)
library(ggrepel)
library(extrafont)
```


# My Coffee Taste Profile
I recorded all of my preferences for all coffees that I have tried. Right now, I'm still getting through two whole bean bags of Peet's coffee, so I won't be able to add in new data for a while.

I put my ratings in `coffee_list.csv`, which you can access <a href="https://github.com/palautatan/yumCoffee/blob/master/datasets/coffee_list.csv">here</a>. Here's a random sample of what is in `coffee_list.csv`.
```{r, include=FALSE, echo=FALSE}
# LOAD IN DATA
my_coffee = read.csv("/Users/edie/Box Sync/GitThings/https---github.com-palautatan-yumCoffee.git/datasets/coffee_list.csv", header=TRUE)
attach(my_coffee)
head(my_coffee)

# CLEAN UP RATING
Rating = unlist(Personal.Rating)
Rating = gsub("\\/5", "", Rating)
Rating = as.numeric(Rating)

# CLEAN UP COFFEE NAMES
coffee_name = str_split(unlist(Coffee),": ")
coffee_name = sapply(coffee_name, function(coffee_part) {
  brand = coffee_part[1]
  name = coffee_part[2]
  c(brand, name)
}
)

# FORMAT DATE
form_date = as.Date(Date, format = "%m/%d/%y")

detach(my_coffee)

## CREATE NEW DATA FRAME
updated_coffee = cbind(my_coffee, Rating, coffee_name[1,], coffee_name[2,], form_date)
names(updated_coffee)[9:11] = c("Brand", "Name", "TrueDate")
attach(updated_coffee)
Roast = factor(Roast, levels=c("Special", "Light", "Medium-Light", "Medium", "French", "Medium-Dark", "Dark"))
```


```{r, echo=FALSE}
srs = sample(length(updated_coffee[,"Coffee"]), 5)
updated_coffee[srs, c("Brand", "Name", "Roast", "Store", "Rating")]
```

I also have some pretty "savage" notes of how I felt about the cofee.

```{r, echo=FALSE, warning=FALSE}
updated_coffee[srs, c("Notes")]
```


# Visualizing Coffee Preferences
Here's a plot of the coffees I have had according to roast. In this plot "Special" means that the coffee was not black.
```{r, echo=FALSE, warning=FALSE}
coffee_theme = theme(panel.background = element_rect(fill = "white", colour = "white"), 
                     panel.grid.minor = element_line(colour = "lightblue", 
                                                     linetype = "dotted"),
                     panel.grid.major = element_line(colour = "lightgrey",
                                                     size = 0.2),
                     plot.title = element_text(size = rel(2), colour = "black",
                                               family="BebasNeue"))
```

### Favorite Roast
Let's look at my favorite roasts! In the following plot, you can see some data points that represent where I got the coffee from, my rating, and the roast of the coffee.

```{r, echo=FALSE, warning=FALSE}
# ROASTS 2
roasts_plot = ggplot()
roasts_plot = roasts_plot + geom_point(aes(x=Roast, y=Rating, colour=factor(Store))) + ylim(0,5)
roasts_plot = roasts_plot + ggtitle("Edie's Roasts")
roasts_plot + coffee_theme
```

This plot shows the same information above, except I labelled the points by store as opposed to coloring them by store.

```{r, echo=FALSE, warning=FALSE}
# ROASTS
coffee_plot = ggplot()
coffee_plot = coffee_plot + geom_point(aes(x=Roast, y=Rating)) + geom_label_repel(aes(x=Roast, y=Rating, label=Brand)) + ylim(0,5)
coffee_plot = coffee_plot + ggtitle("Edie's Roasts")
coffee_plot + coffee_theme
```

I know that many statisticians hate pie charts, but you can at least admit they're pretty.

```{r, echo=FALSE, warning=FALSE}
# PIE CHART
roast_table = table(Roast)
roast_table = data.frame(roast_table)
coffee_pie = ggplot(roast_table, 
                    aes(x="", y=Freq, fill=Roast)) + geom_bar(width = 1, stat = "identity")
coffee_pie = coffee_pie + coord_polar("y", start=0)
coffee_pie = coffee_pie + ggtitle("Edie's Roasts") + xlab("") + ylab("")
coffee_pie + coffee_theme
```

The following calculation tells me what my favorite roasts are according to my taste profile. I don't agree with it though because I know that I really like French Roast. I just haven't had enough French Roasts or had more decent Dark Roasts.

```{r, echo=FALSE}
# What is this taster's favorite roast?
df_roast = sapply(levels(Roast), function(this_level) {
  these_rows = as.numeric(which(Roast==this_level))
  stat = mean(Rating[these_rows])
  c(this_level, stat)
}
)
fave_roast = df_roast[1,which(df_roast[2,]==max(df_roast[2,]))]
cat("This taster's favorite drinks have been from:\n", paste0(fave_roast, sep=". "))
```



### Favorite Store
Here's a plot of where I bought my coffee. The overlapping text in the plot is annoying, yes, but I think it's legible enough. If you cannot read what the first three labels say, here you go: "Costco", "Never Too Latte", and "Ross".

```{r, echo=FALSE, warning=FALSE}
# STORES
coffee_plot_2 = ggplot()
coffee_plot_2 = coffee_plot_2 + geom_point(aes(x=Store, y=Rating, colour=factor(Roast)))
coffee_plot_2 = coffee_plot_2 + ggtitle("Edie's Stores")
coffee_plot_2 + coffee_theme
```

Apparently, my favorite store to buy coffee is Costco. I'm not surprised, and I'm glad my membership is doing me some good.

```{r, echo=FALSE}
# What is this taster's favorite store? (Most frequented)
table_store = rbind(levels(Store), table(Store))
fave_store = table_store[1,which(table_store[2,]==max(table_store[2,]))]
cat("This taster frequently goes to: ", fave_store, ".", sep="")
```


### Top 3 Coffees
Let's look at my Top 3 Coffees over time. Here we see a plot of ratings over time just to examine what kind of bias I'm perpetuating as I rate the coffees I try.

```{r, echo=FALSE, warning=FALSE}
# TIME SERIES OF RATINGS
rate_time = ggplot(data=updated_coffee[which(!is.na(TrueDate)),], aes(x=TrueDate, y=Rating)) + ylim(0,5)
rate_time = rate_time + geom_point(aes(size=Rating, colour=factor(Roast)), alpha=0.75) + scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") 
rate_time = rate_time + ggtitle("Edie's Ratings Over Time") + xlab("Date")
rate_time + coffee_theme + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Here is a density plot of the coffee ratings I gave out. We can obviously see that my ratings come from distinct distributions.

```{r, echo=FALSE, warning=FALSE}
# HISTOGRAM OF SCORES
score_hist = ggplot(updated_coffee[which(Roast!="Special"),], aes(x=Rating, fill=Roast, colour=Roast))
score_hist = score_hist + geom_density(alpha=0.5) + xlim(-1,5)
score_hist = score_hist + ggtitle("Edie's Rating Distributions")
score_hist + coffee_theme
```

My average taste rating is 2.42. I don't think this makes me a coffee prude, nor does it justify my taste profile. I think it means that I really hate Instant Coffee, and several 0's brought down my mean score a lot.

```{r, echo=FALSE}
# What is this taster's average store?
avg_score = mean(Rating)
cat("This taster's average rating is: ", avg_score, ".", sep="")
```

This says that my favorite coffees ever have been San Francisco Bay French Roast OneCup, Joe's Tall Dark and Handsome, and the Pistachio Mocha Latte from Never Too Latte. I think that these are good estimates for my favorites, although I am not actually that much of a fan of super sugary drinks unless they're boba drinks. So, I also calculated my Top 3 without Special drinks included, I would probably do analyses on "Special" drinks and black coffee in different sections.

```{r, echo=FALSE}
# How about top 3 drinks?
# This takes the first 3 sorted regardless of ties
ratings_coffees = rbind(Rating, Name)
ratings_coffees = ratings_coffees[,order(Rating)]
indices = length(ratings_coffees[1,])-c(0:2)
top_three = levels(Name)[ratings_coffees[2, indices]]
cat("Including Special drinks, this taster's top 3 favorite drinks are:\n", paste0(top_three, sep=". "))


# Black coffee only?
ratings_blacks = rbind(updated_coffee[which(Roast!="Special"),]$Rating, updated_coffee[which(Roast!="Special"),]$Name)
ratings_blacks = ratings_blacks[,order(updated_coffee[which(Roast!="Special"),]$Rating)]
indices = length(ratings_blacks[1,])-c(0:2)
top_three_blacks = levels(updated_coffee[which(Roast!="Special"),]$Name)[ratings_blacks[2, indices]]
cat("Excluding special drinks, this taster's top 3 favorite drinks are:\n", paste0(top_three_blacks, sep=". "))
```